---
title: "a3_task3_shane_dewees"
author: "Shane Dewees"
date: "11/10/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(here)
library(kableExtra)
library(lubridate)
library(janitor)
```

## Reading in data

```{r, warning=FALSE}
urchins <- read.csv(here("data", "urchins.csv"))
urchins <- clean_names(urchins)
```

## Exploratory data visualization to become aquainted with the data 
  1. Histogram of size distribution between red and purple urchins. 
  2. Box plot of size vs. common name. 
  
```{r, warning=FALSE}
ggplot(urchins, aes(size, fill = common_name))+
  geom_histogram(binwidth = 0.5)+
  facet_wrap(~common_name)+
  labs(x = "Urchin diameter (cm)", 
       y = "Count")

ggplot(urchins, aes(common_name, size, fill = common_name))+ 
  geom_boxplot() +
  labs(x = "Urchin species",
       y = "Urchin diameter (cm)")
```
  
## Table of Summary statistics to look at how Urchin diameter varies between species. 

```{r, warning=FALSE, message=FALSE}
urchin_stats <- urchins %>% 
  group_by(common_name) %>% 
  summarize(mean = mean(size),
            median = median(size),
            standard_deviation = sd(size),
            sample_size = n())
kable(urchin_stats)
```

When comparing urchin diameter between red and purple urchins, red urchins appear to have larger mean (red urchins = 6.78cm; purple urchins = 4.47cm) and median (red urchins = 7.0cm; purple urchins = 4.5cm) diameter size. Comparing the distribution of diameter sizes, both species appear to have fairly normally distributed data, however, purple urchins seem to have a tighter spread (sd = 1.18cm) compared to red urchins (sd = 1.89cm). Additionally, the percent difference in means (((red-purple)/red)*100) is only 34%, showing that while a difference might be present, it is not that large. 



```{r, warning=FALSE}
purple_urchins <- urchins %>% 
  filter(common_name == "Purple Urchin") %>% 
  mutate(date = mdy(date)) %>% 
  mutate(year = year(date)) %>% 
  mutate(site_full = case_when(
    site == "NAPL" ~"Naples Reef",
    site == "AQUE" ~ "Arroyo Quemado Reef", 
    site == "CARP" ~ "Carpenteria Reef", 
    site == "MOHK" ~ "Mohawk Reef",
    site == "IVEE" ~ "Isla Vista Reef"
  ))
purple_urchins_summary <- purple_urchins %>% 
  group_by(site_full) %>% 
  count(year)

ggplot(purple_urchins_summary, aes(year, n, col = site_full))+
  geom_line() + 
  scale_x_continuous(breaks = seq(from= 2007, to = 2020, by = 3))
```

